---
title: "A2_supp_analysis"
author: "Zoe Woodhead"
date: "5 September 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

`r Sys.time()`

## Background
Original protocol for project is here: <https://osf.io/tkpm2/registrations/>

See Dopp_explore_log.RMd for more background on how Modeling was developed.
See A2_main_analysis.Rmd for analyses reported in the Woodhead et al. (2018) manuscript.


##Chunks of code for setting packages and reading in data 
from SEM2_maxLikelihood_realdata.R
```{r packages, warning=FALSE, message=FALSE}
#Needs OpenMx, which you get with following command (not CRAN)
#Hmm - having problems with OpenMx - now using Cran version

#source('https://openmx.ssri.psu.edu/software/getOpenMx.R')
require(tidyverse)
require(OpenMx)
require(semTools) #for fit measures
require(stringr)
require(data.table)
library('devtools')
library(DiagrammeR)
library(DiagrammeRsvg)
library(magrittr)
library(svglite)
library(rsvg)


########### Set toggles before running! ################################################################
RHonly            <- 1 # Set to 0 if want to include L handers
usemean           <- 1 # Set to 0 to analyse original peak-based LI values
remove.Ntrial     <- 1 # Set to 1 to run exclusion of those with < 12 good trials
tryreorder        <- 1 # Set to 1 to change the fixed path in the SEM model
run_pathdrop      <- 1 # Set to 1 to run pathdrop SEM analysis
run_drawpaths     <- 1 # Set to 1 to run graphviz paths plot

nuorder           <-   # Sets the new order of tasks. The task listed first (and seventh) will be fixed.
# c(3,1,2,4,5,6,9,7,8,10,11,12) # Semantic decision fixed
# c(2,1,3,4,5,6,8,7,9,10,11,12) # Phonological decision fixed
 c(4,1,2,3,5,6,10,7,8,9,11,12) # Sentence generation fixed **PUBLISHED IN MANUSCRIPT**
# c(5,1,2,3,4,6,11,7,8,9,10,12) # Sentence comprehension fixed
# c(6,1,2,3,4,5,12,7,8,9,10,11) # Jabberwocky decision fixed

########################################################################################################

```


##Code for reading and reshaping LI data
N.B. Can select to use original LI values based on peak, or LI values from mean in POI.
This chunk reads in data from .csv files, selects relevant participants, selects peak or mean LI values and identifies excluded data-points

```{r readdata}

# Identify participants to include based on A2_Participant_Info.csv
particdat<-read.csv('A2_Participant_Info.csv')
rhanded<-which(particdat$handedness=='R')
if (RHonly==1){ #See RHonly toggle
  print('R handers only included in analysis')
  particdat <- particdat[rhanded,]
}

# Read in LI data from sessions 1 and 2
data1<-read.csv('Results_Session1.csv')
data2<-read.csv('Results_Session2.csv')

alltaskall <- cbind(select(data1,A1.LI,B1.LI,C1.LI,D1.LI,E1.LI,F1.LI),
             select(data2,A2.LI,B2.LI,C2.LI,D2.LI,E2.LI,F2.LI))

if(usemean==1){ #Select mean LI values rather than peak, if toggled
  alltaskall <- cbind(select(data1,A1.LI_mean,B1.LI_mean,C1.LI_mean,D1.LI_mean,E1.LI_mean,F1.LI_mean),
                      select(data2,A2.LI_mean,B2.LI_mean,C2.LI_mean,D2.LI_mean,E2.LI_mean,F2.LI_mean))
}

mylabels<-c('ListGen1','PhonDec1','SemDec1','SentGen1','SentComp1','Jabber1',
            'ListGen2','PhonDec2','SemDec2','SentGen2','SentComp2','Jabber2')
myshortlab<-c('ListGen','PhonDec','SemDec','SentGen','SentComp','Jabber')

myfixed<-myshortlab[nuorder[1]] ### Keep an eye on this......

colnames(alltaskall)<-mylabels
head(alltaskall)

# Select only right handed participants, if toggled
if (RHonly==1){
  alltaskall<-alltaskall[rhanded,]
}

# Identify datapoints to exclude
alltask<-alltaskall #create a new copy where the outliers will be coded as NA
allse <- cbind(select(data1,A1.se,B1.se,C1.se,D1.se,E1.se,F1.se),
             select(data2,A2.se,B2.se,C2.se,D2.se,E2.se,F2.se))

if (usemean==1){ #See usemean toggle
  allse <- cbind(select(data1,A1.mean_se,B1.mean_se,C1.mean_se,D1.mean_se,E1.mean_se,F1.mean_se),
                 select(data2,A2.mean_se,B2.mean_se,C2.mean_se,D2.mean_se,E2.mean_se,F2.mean_se))
}

if (RHonly==1){ #See RHonly toggle
  allse<-allse[rhanded,]
}

myse<-c(allse[,1],allse[,2],allse[,3],allse[,4],allse[,5],allse[,6],
        allse[,7],allse[,8],allse[,9],allse[,10],allse[,11],allse[,12])

Q3<-quantile(myse,.75)
Q1<-quantile(myse,.25)
Qlimit<-Q3+2.2*(Q3-Q1)
secols<-colnames(allse)

for (i in 1:12){
  w<-which(allse[,i]>Qlimit)
  if (length(w)>0){
    alltask[w,i]<-NA
  }
}

#Now remove those with fewer than 12 trials in a condition
if (remove.Ntrial==1){ #See remove.Ntrial toggle
  nbit<-c('A1.N','B1.N','C1.N','D1.N','E1.N','F1.N')
  includesubs<-1:nrow(data1)
  if(RHonly==1){
    includesubs<-rhanded
  }
  
  for (i in 1:6){
    w<-which(colnames(data1)==nbit[i])
    ww<-which(data1[includesubs,w]<12)
    if(length(ww>0)){
      alltask[ww,i]<-NA
    }
  }
  
  nbit<-c('A2.N','B2.N','C2.N','D2.N','E2.N','F2.N')
  for (i in 1:6){
    w<-which(colnames(data2)==nbit[i])
    ww<-which(data2[includesubs,w]<12)
    if(length(ww>0)){
      alltask[ww,(i+6)]<-NA
    }
  }
}

# Reorder tasks if required
if (tryreorder==1){ #See tryreorder toggle
  alltask<-alltask[,nuorder] #See nuorder toggle
  alltaskall<-alltaskall[,nuorder]
  mylabels<-mylabels[nuorder]
  myshortlab<-myshortlab[nuorder[1:length(myshortlab)]]
}
```


##Run drop-one SEM loop
This chunk drops one subject in each iteration and runs Model 6 (single factor) and Model 7 (bifactor).
```{r dropone}
#Set up model definition

# residual variances
resVars             <- mxPath( from=mylabels, arrows=2,
                        free=c(T,T,T,T,T,T,T,T,T,T,T,T), values=c(1,1,1,1,1,1,1,1,1,1,1,1),
                        labels=c("e1","e2","e3","e4","e5","e6",
                                 "e1","e2","e3","e4","e5","e6"))

# latent variance
latVar_Model6       <- mxPath( from="Factor1", arrows=2,
                        free=T, values=1, labels ="varFactor1" )
latVar_Model7      <- mxPath( from=c("Factor1","Factor2"), arrows=2, connect="unique.pairs",
                        free=c(T,F,F), values=c(1,0,1), labels=c("varFactor1","cov","varFactor2") )

# factor loadings
facLoads_Model6     <- mxPath( from="Factor1", to=mylabels, arrows=1,
                        free=c(FALSE,TRUE,TRUE,TRUE,TRUE,TRUE,
                               FALSE,TRUE,TRUE,TRUE,TRUE,TRUE), #The first path is fixed at one - others scaled relative to this
                        values=c(1,1,1,1,1,1,1,1,1,1,1,1),
                        labels =c("l1","l2","l3","l4","l5","l6",
                                  "l1","l2","l3","l4","l5","l6") )#same for each test on time 1 and 2
facLoadsFactor1_Model7     <- mxPath( from="Factor1", to=mylabels, arrows=1,
                               free=c(FALSE,TRUE,TRUE,TRUE,TRUE,TRUE,FALSE,TRUE,TRUE,TRUE,TRUE,TRUE), 
                               values=rep(1,12),
                               labels =c("k1","k2","k3","k4","k5","k6","k1","k2","k3","k4","k5","k6") )
facLoadsFactor2_Model7     <- mxPath( from="Factor2", to=mylabels, arrows=1,
                               free=c(FALSE,TRUE,TRUE,TRUE,TRUE,TRUE,FALSE,TRUE,TRUE,TRUE,TRUE,TRUE), 
                               values=c(0,rep(1,5),0,rep(1,5)),
                               labels =c("l1","l2","l3","l4","l5","l6","l1","l2","l3","l4","l5","l6") )

# means 
means_Model6        <- mxPath( from="one", to=c(mylabels,'Factor1'), arrows=1,
                        free=c(T,T,T,T,T,T,T,T,T,T,T,T,FALSE), values=c(1,1,1,1,1,1,1,1,1,1,1,1,0),
                        labels =c(myshortlab, myshortlab, NA) ) 
                        #means constant from time 1 to time 2. 
                        #One extra mean for the Factor, but this is set to NA
means_Model7        <- mxPath( from="one", to=c(mylabels,'Factor1','Factor2'), arrows=1,
                        free=c(T,T,T,T,T,T,T,T,T,T,T,T,FALSE,FALSE), values=c(1,1,1,1,1,1,1,1,1,1,1,1,0,0),
                        labels =c(myshortlab, myshortlab, NA,NA) )

#Create dataframe to save the results
nsubs<-nrow(alltaskall)
dropone_results <- matrix(data = NA, nrow = nsubs, ncol = 5)
  
#Use'drop one' approach - reduce the dataset by one on each run

for (thisdrop in 1:nsubs){
  dataRaw      <- mxData( observed=alltask[-thisdrop,], type="raw" )
  
  
  #Estimate Model 6
  myModel6 <- mxModel("Single Factor Model", type="RAM",
                          manifestVars=mylabels, latentVars="Factor1",
                          dataRaw, resVars, latVar_Model6, facLoads_Model6, means_Model6)
  Model6Fit <- mxRun(myModel6)
  summary.Model6<-summary(Model6Fit, refModels=mxRefModels(Model6Fit, run = TRUE))   
  
  #Estimate Model 7
  myModel7 <- mxModel("Bifactor Model", type="RAM",
                          manifestVars=mylabels, latentVars=c("Factor1", "Factor2"),
                          dataRaw, resVars, latVar_Model7, facLoadsFactor1_Model7, facLoadsFactor2_Model7, means_Model7) 
  Model7Fit <- mxRun(myModel7)
  summary.Model7<-summary(Model7Fit, refModels=mxRefModels(Model7Fit, run = TRUE)) 
  
  #Compare Models
  mcomp<-mxCompare(Model7Fit, Model6Fit)
  dropone_results[thisdrop, 1] <- round(summary.Model6$BIC.Mx, 3)
  dropone_results[thisdrop, 2] <- round(summary.Model7$BIC.Mx, 3)
  dropone_results[thisdrop, 3] <- round(mcomp$diffLL[2], 3)
  dropone_results[thisdrop, 4] <- mcomp$diffdf[2]
  dropone_results[thisdrop, 5] <- round(mcomp$p[2], 3)

}


colnames(dropone_results) <- c('BIC_mod6','BIC_mod7','chi.diff','df.diff','p.diff')
dropone_results <- cbind(dropone_results, dropone_results[ , 1] - dropone_results[ ,2])
Mod6_wins <- length(which(dropone_results[,5] < .05 & dropone_results[,6] > 0))
Mod7_wins <- length(which(dropone_results[,5] < .05 & dropone_results[,6] < 0))

mod6_message <- paste0('Model 6 wins in ', Mod6_wins, ' iterations out of 30')
mod7_message <- paste0('Model 7 wins in ', Mod7_wins, ' iterations out of 30')

```